# Reproducible Research: Peer Assessment 1


### Loading and preprocessing the data
```{r this section unzips and upload the data}
unzip('activity.zip')
activity<-read.csv('activity.csv', colClasses=c("integer","Date","factor"))
activity$month <-as.numeric(format(activity$date,"%m"))
## excludes NA values
activity_noNA<-na.omit(activity)
rownames(activity_noNA)<-1:nrow(activity_noNA)
library(ggplot2)
```

### What is mean total number of steps taken per day?

#### 1.Make a histogram of the number of steps taken each day

```{r makes the plot and calculates the mean}
## makes the plot
ggplot(activity_noNA, aes(date, steps)) + geom_bar(stat = "identity", colour = "green" , fill = "green", width = 0.8) + facet_grid(. ~ month, scales = "free") + labs(title = "1.Histogram of Total Number of Steps Taken Each Day", x = "Date", y = "Number of steps")
```

#### 2. Calculate and report the <b>mean</b> and median total number of steps taken per day

```{r processes the database to calculate the means}
## it is needed to use a funcion to sum the total number of steps taken each day
totalstepsperday<-aggregate(activity_noNA$steps, list(Date=activity_noNA$date), FUN = "sum")$x
```
Mean
```{r, echo=FALSE}
mean(totalstepsperday)
```
Median
```{r, echo=FALSE}
median(totalstepsperday)
```



## What is the average daily activity pattern?



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
