# Reproducible Research: Peer Assessment 1


### Loading and preprocessing the data
```{r this section unzips and upload the data}
unzip('activity.zip')
activity<-read.csv('activity.csv', colClasses=c("integer","Date","factor"))
activity$month <-as.numeric(format(activity$date,"%m"))
## excludes NA values
activity_noNA<-na.omit(activity)
rownames(activity_noNA)<-1:nrow(activity_noNA)
library(ggplot2)
```

### What is mean total number of steps taken per day?

#### 1. Make a histogram of the number of steps taken each day

```{r makes the plot and calculates the mean}
## makes the plot
ggplot(activity_noNA, aes(date, steps)) + geom_bar(stat = "identity", colour = "green" , fill = "green", width = 0.8) + facet_grid(. ~ month, scales = "free") + labs(title = "1.Histogram of Total Number of Steps Taken Each Day", x = "Date", y = "Number of steps")
```

#### 2. Calculate and report the <i>mean</i> and <i>median</i> total number of steps taken per day

```{r processes the database to calculate the means}
## it is needed to use a funcion to sum the total number of steps taken each day
totalstepsperday<-aggregate(activity_noNA$steps, list(Date=activity_noNA$date), FUN = "sum")$x
```
Mean
```{r calculate mean}
mean(totalstepsperday)
```
Median
```{r calculate median}
median(totalstepsperday)
```



### What is the average daily activity pattern?

#### 1. Make a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days(y-axis)

```{r Average steps by interval}
stepsbyinterval<- aggregate(activity_noNA$steps, list(interval = as.numeric(as.character(activity_noNA$interval))), FUN = "mean")
plot(stepsbyinterval$interval,stepsbyinterval$x,type='l',main="Average steps by 5-minute interval",xlab="5-minute interval",ylab="average steps")
```

#### 2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r 5-minute interval with more steps}
stepsbyinterval[stepsbyinterval$x==max(stepsbyinterval$x),]$interval
```

## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
